---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
After forking and cloning [this](https://github.com/rdpeng/RepData_PeerAssessment1) repository, set the working directory in R.   
```{r eval=FALSE}
setwd("./RepData_PeerAssessment1") 
```
Load the data into R:
```{r}
activity <- read.csv(file=unzip("activity.zip"))  
activity$date <- as.Date(as.character(activity$date, "%Y-%m-%d"))
```
## What is mean total number of steps taken per day?
The total number of steps taken per day:
```{r}
with(activity, tapply(steps, date, sum, na.rm=T))
```
A histogram of the total number of steps taken per day:
```{r histogram}
sumsteps <- aggregate(steps~date, data=activity, sum)
hist(sumsteps$steps, col="red", main="Total Number of Steps per Day",
     xlab="Steps per Day")   
```   

The mean and median of the total number of steps taken per day:
```{r}
mean(sumsteps$steps)
median(sumsteps$steps)
```
## What is the average daily activity pattern?
Calculation of the average number of steps per interval followed by a time series plot of the 5-minute interval and the average number of steps:  
```{r time series plot}
meansteps <- aggregate(steps~interval, data=activity, mean) 
with(meansteps, plot(interval, steps, type="l", 
    main="Average Number of Steps Taken", ylab="Average steps"))   
```   

The 5-minute interval with the maximum number of steps:
```{r}
meansteps[which.max(meansteps$steps), 1]  
```
## Imputing missing values
Total number of NA values:
```{r}
sum(is.na(activity))
```
Imputing the NA values with the median of the steps per interval:
```{r}
mediansteps <- aggregate(steps~interval, data=activity, median)
missing <- activity[is.na(activity$steps), ]
imputed1 <- merge(missing, mediansteps, by="interval")
imputed1 <- imputed1[order(imputed1$date), ]
activity1 <- activity
activity1[is.na(activity1$steps), 1] <- imputed1$steps.y
```
Histogram of the total number of steps taken each day from imputed data:
```{r histogram with imputed data}
newsumsteps <- aggregate(steps~date, data=activity1, sum)
with(newsumsteps, hist(steps, col="red"))  
```  

The mean and median of the total number of steps taken per day:
```{r}
mean(newsumsteps$steps)
median(newsumsteps$steps)
```
Imputing the data caused the mean and median to both decrease.  

The total number of steps per day using the imputed data:
```{r}
with(newsumsteps, tapply(steps, date, sum, na.rm=T))
```
## Are there differences in activity patterns between weekdays and weekends?
Create the new column with factor variable with levels weekend and weekday:
```{r}
activity2 <- activity1
activity2$week <- ifelse(weekdays(activity2$date)=="Saturday" | 
            weekdays(activity2$date)=="Sunday", "weekend", "weekday")
activity2$week <- as.factor(activity2$week)
```
A time series panel plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days and weekend days:
```{r panel plot}
meansteps2 <- aggregate(steps~interval + week, data=activity2, mean)
library(lattice)
xyplot(steps~interval | week, data=meansteps2, type = "l", layout=c(1, 2))
```